---
title: "Covid-19 in numbers"
subtitle: "A data analysis using publicly available data"
author: "Dr. Jörg Padberg"
output: 
  html_notebook:
    code_folding: hide
---

# {.tabset}

```{r, message=F}
# Load packages
library(OECD)
library(zoo)
library(readr)
library(stringr)
library(dplyr)
library(tidyr)
library(tsibble)
library(ggplot2)
library(ggpubr)

library(cartography)
library(sp)

library(eurostat) # for more up-to-date map data for Germany

# Define custom colors
eco_col <- "olivedrab3"

# Define population
pop_de <- 83.1*10^6
```


## Germany

```{r, message=F}
# Germany
# rki <- read_csv("https://opendata.arcgis.com/datasets/dd4580c810204019a7b8eb3e0b329dd6_0.csv")
# write_csv(rki, "covid19_rki.csv")
rki <- read_csv("covid19_rki.csv")

# Prepare data
rki$Geschlecht <- factor(rki$Geschlecht)
rki$Meldedatum <- as.Date(rki$Meldedatum)
rki$Datenstand <- as.Date(rki$Datenstand, format = "%d.%m.%Y")
rki$Refdatum <- as.Date(rki$Refdatum)

```

Show a map. The current situation first, then the history...

```{r, message=F, fig.asp=0.8}
rki_counties <- read_csv("https://opendata.arcgis.com/datasets/917fc37a709542548cc3be077a786c17_0.csv")

# Get a more up-to-date spatial polygons data frame of Germany
map_de2 <- get_eurostat_geospatial(output_class = "spdf",
                                   resolution = "60",
                                   nuts_level = "3",
                                   year = "2016")

# Calculate average incidence for Berlin (otherwise not in map)
rki_berlin <- rki_counties %>% 
  filter(str_detect(GEN, "Berlin")) %>% 
  select(GEN, BEZ, cases7_per_100k, NUTS)

avg_berlin <- mean(rki_berlin$cases7_per_100k)

rki_counties_with_berlin <- rki_counties %>% 
  select(GEN, cases7_per_100k, NUTS) %>% 
  add_row(GEN = "Berlin", cases7_per_100k = avg_berlin, NUTS = "DE300")

# Plot the choropleth map
par(mar = c(0,1,2,1))
choroLayer(spdf = map_de2, 
           df = rki_counties_with_berlin, 
           dfid = "NUTS", 
           var = "cases7_per_100k", 
           col = carto.pal("red.pal", n1 = 6),
           method = "q6",
           legend.title.txt = "Cases per 100.000",
           legend.pos = "right",
           legend.frame = T)
layoutLayer(title = "Confirmed Covid-19 cases in Germany per county (7-day incidence)",
            frame = F,
            col = "white",
            coltitle = "black",
            author = paste("Data in this map from:", unique(rki_counties$last_update)),
            north = T,
            scale = F) # because of unrealistic number
```

```{r, fig.width=10, fig.asp=0.5}
# History as a line-chart or as a faceted choropleth map? -> line-chart first
rki %>% 
  select(Meldedatum, AnzahlFall, Landkreis) %>% 
  group_by(Meldedatum) %>% 
  summarise(new_cases = sum(AnzahlFall)) %>% 
  mutate(incidence_7day = rollmean(new_cases, 7, fill = NA)/pop_de*10^5*7) %>% 
  ggplot(aes(x = Meldedatum, y = incidence_7day)) +
    geom_line() +
    geom_area(alpha = 0.3) +
    geom_hline(yintercept = 50, lty = 2) +
    theme_classic() +
    labs(title = "Development of the Covid-19 7-day incidence in Germany",
         x = "Datum", 
         y = "7-day incidence")
```



```{r, message=F, warning=F, fig.width=10}
cases_de_relative <- rki %>% 
  group_by(Meldedatum) %>% 
  summarise(new_cases = sum(AnzahlFall)) %>% 
  mutate(new_cases_100k = 10^5*new_cases/pop_de,
         rollmean_cases_100k = rollmean(new_cases_100k, 7, fill = NA)) %>% 
  ggplot(aes(x = Meldedatum, y = new_cases_100k)) +
    geom_line(lwd = 0.2) +
    geom_line(aes(y = rollmean_cases_100k), lwd = 1) +
    theme_classic() +
    labs(title = "Confirmed Covid-19 cases in Germany",
         subtitle = "Cases per 100.000 population")

cases_de_absolute <- rki %>% 
  group_by(Meldedatum) %>% 
  summarise(new_cases = sum(AnzahlFall)) %>% 
  mutate(rollmean_cases = rollmean(new_cases, 7, fill = NA)) %>% 
  ggplot(aes(x = Meldedatum, y = new_cases)) +
    geom_line(lwd = 0.2) +
    geom_line(aes(y = rollmean_cases), lwd = 1) +
    theme_classic() +
    labs(subtitle = "Absolute number of confirmed cases")

deaths_de <- rki %>% 
  group_by(Meldedatum) %>% 
  summarise(deaths = sum(AnzahlTodesfall)) %>% 
  mutate(rollmean_deaths = rollmean(deaths, 7, fill = NA)) %>% 
  ggplot(aes(x = Meldedatum, y = deaths)) +
    geom_line(lwd = 0.2) +
    geom_line(aes(y = rollmean_deaths), lwd = 1) +
    theme_classic() +
    labs(subtitle = "Deaths by or in the presence of Covid-19")

nest_de <- ggarrange(cases_de_absolute, deaths_de, nrow = 1)

ggarrange(cases_de_relative, nest_de, ncol = 1, heights = c(2,1))
```

How old are the patients?

```{r, message=F, fig.width=10, fig.asp=0.4}
rki_age_classes <- rki %>% 
  filter(Geschlecht != "unbekannt") %>% 
  group_by(Altersgruppe, Geschlecht) %>% 
  summarise(confirmed_cases = sum(AnzahlFall),
            deaths = sum(AnzahlTodesfall)) %>% 
  gather(confirmed_cases, deaths, key = "case", value = "number") 

rki_age_classes %>% 
  ggplot(aes(x = Altersgruppe, y = number, fill = Geschlecht)) +
    geom_col(position = "dodge", alpha = 0.7) +
    scale_y_continuous(labels = scales::comma) +
    scale_fill_brewer(palette = "Set1", direction = -1) +
    facet_wrap(~case, scales = "free_y") +
    theme_classic()
```

How does the population pyramid looks like for the same age classes?

```{r, message=F}
population_pyramid_de <- read_csv("demographic_de_2018.csv")

population_pyramid_de %>% 
  ggplot(aes(x = Altersgruppe, y = Bevölkerung, fill = Geschlecht)) +
    geom_col(position = "dodge", alpha = 0.7) +
    scale_fill_brewer(palette = "Set1", direction = -1) +
    theme_classic() +
    labs(y = "Bevölkerung (in Tausend)")
```

Now a diagram showing, how big the share in each age-class is...

```{r, message=F, fig.width=10, fig.asp=0.4}
# Bring the data together
left_join(rki_age_classes, population_pyramid_de) %>% 
# Calculate the share of Covid-19 cases
  mutate(cases_pct = number*100/(Bevölkerung*1000)) %>% 
# Visualisation
  filter(Altersgruppe != "unbekannt") %>% 
  ggplot(aes(x = Altersgruppe, y = cases_pct, fill = Geschlecht)) +
    geom_col(position = "dodge", alpha = 0.7) +
    scale_fill_brewer(palette = "Set1", direction = -1) +
    facet_wrap(~case, scales = "free_y") +
    theme_classic()
  
```


... maybe a stacked area chart to see how the age classes change over the pandemic?

```{r, message=F, fig.width=10, fig.asp=0.4}
rki %>% 
  filter(Geschlecht != "unbekannt") %>% 
  group_by(Meldedatum, Geschlecht) %>% 
  summarise(confirmed_cases = sum(AnzahlFall),
            deaths = sum(AnzahlTodesfall)) %>% 
  ungroup() %>% 
  mutate('7day_avg_cases' = zoo::rollmean(confirmed_cases, 7, fill = NA),
         '7day_avg_deaths' = zoo::rollmean(deaths, 7, fill = NA)) %>% 
#  gather(confirmed_cases, deaths, key = "type", value = "number") %>% 
  select(-confirmed_cases, -deaths) %>% 
  gather('7day_avg_cases', '7day_avg_deaths', key = "type", value = "number") %>% 
  ggplot(aes(x = Meldedatum, y = number, fill = Geschlecht)) +
    geom_area(alpha = 0.7, position = "fill") +
    facet_wrap(~type) +
    scale_fill_brewer(palette = "Set1", direction = -1) +
    theme_classic() +
    labs(title = "Gender relation of confirmed Covid-19 cases and deaths",
         x = NULL,
         y = "Share")
```


... what about detailed data regarding ortenau? vs. germany?

```{r, message=F}
pop_ortenau <- 430953

rki %>% 
  filter(Landkreis == "LK Ortenaukreis") %>% 
  group_by(Meldedatum) %>% 
  summarise(new_cases = sum(AnzahlFall)) %>% 
  mutate(new_cases_100k = 10^5*new_cases/pop_ortenau,
         rollmean_cases_100k = rollmean(new_cases_100k, 7, fill = NA)) %>% 
  ggplot(aes(x = Meldedatum, y = new_cases_100k)) +
    geom_line(lwd = 0.2) +
    geom_line(aes(y = rollmean_cases_100k), lwd = 1) +
    theme_classic() +
    labs(title = "Confirmed Covid-19 cases Ortenau County",
         subtitle = "Cases per 100.000 population")
```


In order to take countermeasures against the economic consequences of the shutdown, the German government as well as other governsments has ... huge amounts of money. This, of course, has an effect on the inflation rate. The following graph illustrates the monthly customer Price Index (CPI) for Germany and the three largest economies of the world for the last five years. 

... all items without seasonal adjustment. 

```{r, fig.width=10}
# ---------------------
# Inflation
# ---------------------

# Import data
#CPI_datasets <- search_dataset("CPI")

#CPI_data_structure <- get_data_structure("PRICES_CPI")

filter_list <- list(NULL, # the location/country
                    "CPALTT01", # the basket (all items in this case)
                    "IXOB", # the kind of index (standard index in this case - no seasonal adjustment)
                    "M") # monthly data

CPI_data <- get_dataset(dataset = "PRICES_CPI",
                        filter = filter_list,
                        start_time = 2015,
                        end_time = 2020)

# Prepare data
CPI_data$obsTime <- as.yearmon(CPI_data$obsTime)

# Visualise the data
CPI_DE <- CPI_data %>% 
  filter(LOCATION == "DEU") %>% 
  ggplot(aes(x = obsTime, y = obsValue)) +
    geom_line() +
    scale_y_continuous(limits = c(NA, 115)) +
    theme_classic() +
    labs(title = "Customer Price Index (CPI)",
         subtitle = "Germany",
         x = "Year",
         y = "CPI")

CPI_EU <- CPI_data %>% 
  filter(LOCATION == "EU27_2020") %>% 
  ggplot(aes(x = obsTime, y = obsValue)) +
    geom_line() +
    scale_x_yearmon(format = "%Y") +
    scale_y_continuous(limits = c(NA, 115)) +
    theme_classic() +
    labs(subtitle = "European Union (27)",
         x = "Year",
         y = "CPI")

CPI_US <- CPI_data %>% 
  filter(LOCATION == "USA") %>% 
  ggplot(aes(x = obsTime, y = obsValue)) +
    geom_line() +
    scale_x_yearmon(format = "%Y") +
    scale_y_continuous(limits = c(NA, 115)) +
    theme_classic() +
    labs(subtitle = "United States",
         x = "Year",
         y = "CPI")

CPI_CN <- CPI_data %>% 
  filter(LOCATION == "CHN") %>% 
  ggplot(aes(x = obsTime, y = obsValue)) +
    geom_line() +
    scale_x_yearmon(format = "%Y") +
    scale_y_continuous(limits = c(NA, 115)) +
    theme_classic() +
    labs(subtitle = "China",
         x = "Year",
         y = "CPI")

nest_CPI <- ggarrange(CPI_EU, CPI_US, CPI_CN, ncol = 1)

ggarrange(CPI_DE, nest_CPI, ncol = 2, widths = c(2,1))
```

How was the effect? Take a look at the GDP...

## EU

```{r}
# Import data
covid19_testing <- read.csv("https://opendata.ecdc.europa.eu/covid19/testing/csv", 
                    na.strings = "", 
                    fileEncoding = "UTF-8-BOM")

# Prepare data
covid19_testing$year <- as.numeric(str_sub(covid19_testing$year_week, start = 1, end = 4))
covid19_testing$week <- as.numeric(str_sub(covid19_testing$year_week, start = 7, end = 8))

covid19_testing$yearweek <- yearweek(as.character(covid19_testing$year_week))

casecol <- "firebrick2"
poscol <- "olivedrab3"
testcol <- "dodgerblue3"
```

The following map provides an overview of the magnitude of confirmed Covid-19 infections in the European Union. Since the European Centre for Disease Prevention and Control (ECDC) only provides weekly data as of December 17th 2020, the map shows the 7-day incidence, i.e. the sum of all confirmed cases in the last calendar week per 100.000 persons. 

```{r, fig.asp=1}
# Prepare the data (only the most recent week is of interest)
covid19_eu <- covid19_testing %>% 
  group_by(country) %>% 
  top_n(1, year_week) %>% 
  ungroup() %>% 
  mutate(new_cases_100k = 10^5*new_cases/population)

# Now the spatial data (Covid-19 data only on the country level)
data(nuts2006)

last_update <- unique(covid19_eu$year_week)

par(mar = c(0,1,2,1))
plot(nuts0.spdf)
choroLayer(spdf = nuts0.spdf, 
           df = covid19_eu, 
           dfid = "country_code", 
           var = "new_cases_100k",
           col = carto.pal("red.pal", n1 = 7),
           method = "pretty",
           legend.title.txt = "Cases per 100.000",
           legend.pos = "right",
           legend.frame = T,
           border = "black",
           add = T)
layoutLayer(title = "Confirmed Covid-19 cases in the EU (7-day incidence)",
            frame = F,
            col = "white",
            coltitle = "black",
            author = paste("Data in this map from:", last_update),
            north = T,
            scale = 500)
labelLayer(spdf = nuts0.spdf, txt = "id", halo = T, show.lines = T)
```


The previous map, however, only shows the current situation without any information regarding the development of the epidemic over time. For this reason, the next diagram illustrates the data of confirmed Covid-19 cases over time since the beginning of the pandemic. The underlying data is again weekly data from the ECDC. This means that each data point represents the number of confirmed infections with Covid-19 in one week. For a better comparison of the different countries, the number of cases was normalised on 100.000 persons. 

```{r, fig.width=10}
# New cases
covid19_testing %>% 
  mutate(new_cases_100k = 10^5*new_cases/population) %>% 
  ggplot(aes(x = week, y = new_cases_100k)) +
    geom_line(color = casecol, lwd = 1) +
    geom_area(fill = casecol, alpha = 0.3) +
    facet_wrap(~country) +
    theme_classic() +
    labs(title = "Weekly number of newly reported Covid-19 cases",
         x = "Week", 
         y = "Confirmed new cases per 100.000 population")

```

Since some people claim that the cases are only that high because we are testing more than in the beginning of the pandemic, the following two diagrams illustrate the testing rate, i.e. the number of performed tests per 100.000 people in the given period (in this case one week) as well as the positivity rate. The latter one is the percentage of positive tests from all performed tests for the given period. 

```{r, fig.width=10}
# Testing rate
covid19_testing %>% 
  ggplot(aes(x = week, y = testing_rate)) +
    geom_line(color = testcol, lwd = 1) +
    geom_area(fill = testcol, alpha = 0.3) +
    facet_wrap(~country) +
    theme_classic() +
    labs(title = "Testing rate",
         x = "Week",
         y = "Testing rate per 100.000 population")

# Positivity rate
covid19_testing %>% 
  ggplot(aes(x = week, y = positivity_rate)) +
    geom_line(color = poscol, lwd = 1) +
    geom_area(fill = poscol, alpha = 0.3) +
    facet_wrap(~country) +
    theme_classic() +
    labs(title = "Positivity rate",
         x = "Week",
         y = "Share of positive tests [%]")

```


## USA

The data is taken directly from the [GitHub repository of the Johns Hopkins University](https://github.com/CSSEGISandData/COVID-19/blob/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv). The following table provides an overview of the data structure using the first six rows of the data set: 

```{r, message=F}
# Daten einlesen
url <- "https://github.com/CSSEGISandData/COVID-19/raw/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv"

covid19_usa <- read_csv(url)

head(covid19_usa)
```

```{r}
covid19_usa_tidy <- covid19_usa %>% 
  select(-c(UID, iso2, iso3, code3, Admin2, Country_Region, Lat, Long_)) %>% 
  gather(-c(FIPS, Province_State, Combined_Key), key = "Date", value = "confirmed_cases") %>% 
  mutate(Date = as.Date(Date, format = "%m/%d/%y"),
         fips = FIPS) %>% # schreibweise wichtig für usmap-Paket
  select(Date, Combined_Key, Province_State, fips, confirmed_cases)

covid19_prepared <- covid19_usa_tidy %>% 
  group_by(Combined_Key) %>% # fips hat NA's. Stattdessen Combined_Key verwenden!
  mutate(previous_day = lag(confirmed_cases, n = 1L),
         new_cases = confirmed_cases-previous_day) %>% 
  ungroup()

# Prepare the countypop data frame
countypop2015 <- usmap::countypop %>% 
  mutate(fips = as.numeric(fips))

# Join the two data frames, exclude redundant variables and calculate cases per 100.000 people
covid19_normalised <- covid19_prepared %>% 
  left_join(countypop2015, by = "fips") %>% 
  select(-abbr, -county) %>% 
  mutate(new_cases_100k = 10^5*new_cases/pop_2015)
```

```{r, message=F, fig.width=10, fig.asp=0.5}
# Total cases for the US
covid19_prepared %>% 
  group_by(Date) %>% 
  summarise(new_cases_total = sum(new_cases)) %>% 
  mutate(rollmean_cases = rollmean(new_cases_total, 7, fill = NA)) %>% 
  ggplot(aes(x = Date, y = new_cases_total)) +
    geom_line(lwd = 0.2) +
    geom_line(aes(y = rollmean_cases), lwd = 1) +
    theme_classic() +
    labs(title = "Confirmed Covid-19 infections in the US",
         x = NULL,
         y = "confirmed new cases per day")
```


```{r, message=F, fig.width=10, fig.asp=1}
# Cases per US-State
covid19_normalised %>% 
  group_by(Date, Province_State) %>% 
  summarise(state_sum = sum(new_cases, na.rm = T)) %>% 
  ggplot(aes(x = Date, y = state_sum)) +
    geom_line() +
    facet_wrap(~Province_State) +
    scale_x_date(date_labels = "%b") +
    theme_classic() +
    labs(title = "Confirmed Covid-19 infections in the US per State",
         x = NULL,
         y = NULL)
```

